---
title: "Performance on incremental size input for Shapes generation"
author: "Herminio García-González"
date: "2024-07-05"
output: 
  html_document: default
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(summarytools)
library(FSA)
library(ggplot2)
library(rstatix)
library(rmarkdown)
library(ARTool)
library(emmeans)
library(car)
library(ggpubr)
library(lmtest)
```

# Set up
```{r}
dataset <- read.csv("C:/Users/Herminio/Git/shapes-generation-evaluation/statistics/resultEvaluationIncrementalFilms.csv", sep=';')[, 2:5]
datasetShEx <- dataset[(dataset$ShapesLanguage == "ShEx"), ]
datasetSHACL <- dataset[(dataset$ShapesLanguage == "SHACL"), ]
```

# Descriptive statistics for SHACL generation
```{r}
stby(datasetSHACL, list(datasetSHACL$Tool, datasetSHACL$Size), descr, round.digits=5, stats=c("mean", "med", "sd", "min", "max"))
```

# Descriptive statistics for ShEx generation
```{r}
stby(datasetShEx, list(datasetShEx$Tool, datasetShEx$Size), descr, round.digits=5, stats=c("mean", "med", "sd", "min", "max"))
```

# Testing normality for SHACL
```{r}
by(datasetSHACL$ElapsedTime_ms, list(datasetSHACL$Tool, datasetSHACL$Size), shapiro.test)
```

# Testing normality for ShEx
```{r}
by(datasetShEx$ElapsedTime_ms, list(datasetShEx$Tool, datasetShEx$Size), shapiro.test)
```


# Running the ART tool including post hoc tests and effect sizes for SHACL
```{r}
datasetSHACL$Size <- factor(datasetSHACL$Size)
datasetSHACL$Tool <- factor(datasetSHACL$Tool)

model = art(ElapsedTime_ms ~ Tool + (1 | Size), data=datasetSHACL)
marginal = art.con(model, "Tool", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `F` * `Df` / (`F` * `Df` + `Df.res`))
Result
marginal

lm = artlm(model, "Tool")
contrasts = summary(pairs(emmeans(lm, ~ Tool)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Comparing by Elapsed Time and Size using the ART tool on incremental SHACL generation
```{r}
options(max.print=999999)

datasetSHACL$Size <- factor(datasetSHACL$Size)
datasetSHACL$Tool <- factor(datasetSHACL$Tool)

model = art(ElapsedTime_ms ~ Tool + Size + Tool:Size, data=datasetSHACL)
marginal = art.con(model, "Tool:Size", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))
Result

lm = artlm(model, "Tool:Size")
contrasts = summary(pairs(emmeans(lm, ~ Tool:Size)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts

# This is to ensure that we can use the Cohen's D calculations as the homoscedasticity of the linear model is guaranteed
bptest(lm)
```

# Unstandardised effect sizes 
```{r}
by(datasetSHACL, datasetSHACL$Size, function(x) {
  comparison <- c()
  diff_means <- c()
  for (i in unique(x$Tool)) {
    m1 <- mean(x[(x$Tool == i), ]$ElapsedTime_ms)
    for (j in unique(x$Tool)) {
      if(i != j) {
        m2 <- mean(x[(x$Tool == j), ]$ElapsedTime_ms)
        if(m2 - m1 > 0) { # this avoids including the same comparison twice
          comparison <- append(comparison, paste(i, "-", j))
          diff_means <- append(diff_means, m2 - m1)
        }
      }
    }
  }
  data.frame(comparison, diff_means)
})
```


# Comparing ShExML and sheXer by Elapsed Time and Size using the ART tool on incremental SHACL generation
```{r}
datasetSHACL$Size <- factor(datasetSHACL$Size)
datasetSHACL$Tool <- factor(datasetSHACL$Tool)

model = art(ElapsedTime_ms ~ Tool + Size + Tool:Size, data=datasetSHACL[(datasetSHACL$Tool == "ShExML" | datasetSHACL$Tool == "sheXer"), ])
marginal = art.con(model, "Tool:Size", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))
Result

lm = artlm(model, "Tool:Size")
contrasts = summary(pairs(emmeans(lm, ~ Tool:Size)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Running the ART tool including post hoc tests and effect size for ShEx
```{r}
datasetShEx$Size <- factor(datasetShEx$Size)
datasetShEx$Tool <- factor(datasetShEx$Tool)

model = art(ElapsedTime_ms ~ Tool + (1 | Size), data=datasetShEx)
marginal = art.con(model, "Tool", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `F` * `Df` / (`F` * `Df` + `Df.res`))
Result
marginal

lm = artlm(model, "Tool")
contrasts = summary(pairs(emmeans(lm, ~ Tool)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Comparing ShExML and sheXer by Elapsed Time and Size using the ART tool on incremental ShEx generation
```{r}
datasetShEx$Size <- factor(datasetShEx$Size)
datasetShEx$Tool <- factor(datasetShEx$Tool)

model = art(ElapsedTime_ms ~ Tool + Size + Tool:Size, data=datasetShEx)
marginal = art.con(model, "Tool:Size", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))
Result

lm = artlm(model, "Tool:Size")
contrasts = summary(pairs(emmeans(lm, ~ Tool:Size)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts

# This is to ensure that we can use the Cohen's D calculations as the homoscedasticity of the linear model is guaranteed
bptest(lm)
```

# Unstandardised effect sizes 
```{r}
by(datasetShEx, datasetShEx$Size, function(x) {
  comparison <- c()
  diff_means <- c()
  for (i in unique(x$Tool)) {
    m1 <- mean(x[(x$Tool == i), ]$ElapsedTime_ms)
    for (j in unique(x$Tool)) {
      if(i != j) {
        m2 <- mean(x[(x$Tool == j), ]$ElapsedTime_ms)
        if(m2 - m1 > 0) { # this avoids including the same comparison twice
          comparison <- append(comparison, paste(i, "-", j))
          diff_means <- append(diff_means, m2 - m1)
        }
      }
    }
  }
  data.frame(comparison, diff_means)
})
```

# Comparing linear scale on SHACL
```{r}
ggline(datasetSHACL, x = "Size",
      y = "ElapsedTime_ms",
      color = "Tool",
      add = "mean_sd")
```

# Comparing logarithmic scale on SHACL
```{r}
ggline(datasetSHACL, x = "Size",
                y = "ElapsedTime_ms",
                ylab = "Elapsed Time (ms)",
                yscale = "log10",
                color = "Tool",
                add = c("mean_se"))
```

# Comparing linear scale on ShEx
```{r}
ggline(datasetShEx, x = "Size",
      y = "ElapsedTime_ms",
      color = "Tool",
      add = "mean_sd")
```

# Comparing logarithmic scale on ShEx
```{r}
ggline(datasetShEx, x = "Size",
      y = "ElapsedTime_ms",
      ylab = "Elapsed Time (ms)",
      yscale = "log10",
      color = "Tool",
      add = "mean_sd")
```

# Comparing ShEx and SHACL generation times on sheXer
```{r}
datasetShexer <- dataset[(dataset$Tool == "sheXer"), ]
datasetShexer$Size <- factor(datasetShexer$Size)
datasetShexer$ShapesLanguage <- factor(datasetShexer$ShapesLanguage)

model = art(ElapsedTime_ms ~ ShapesLanguage + (1 | Size), data=datasetShexer)
marginal = art.con(model, "ShapesLanguage", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `F` * `Df` / (`F` * `Df` + `Df.res`))
Result
marginal

lm = artlm(model, "ShapesLanguage")
contrasts = summary(pairs(emmeans(lm, ~ ShapesLanguage)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```
# Running the ART tool including post hoc tests and effect size for sheXer
```{r}
model = art(ElapsedTime_ms ~ ShapesLanguage + Size + ShapesLanguage:Size, data=datasetShexer)
marginal = art.con(model, "ShapesLanguage:Size", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))
Result

lm = artlm(model, "ShapesLanguage:Size")
contrasts = summary(pairs(emmeans(lm, ~ ShapesLanguage:Size)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Comparing linear scale on sheXer
```{r}
ggline(datasetShexer, x = "Size",
      y = "ElapsedTime_ms",
      color = "ShapesLanguage",
      add = "mean_sd")
```

# Comparing logarithmic scale on sheXer
```{r}
ggline(datasetShexer, x = "Size",
      y = "ElapsedTime_ms",
      ylab = "Elapsed Time (ms)",
      yscale = "log10",
      color = "ShapesLanguage",
      add = "mean_sd")
```

# Comparing ShEx and SHACL generation times on ShExML
```{r}
datasetShExML <- dataset[(dataset$Tool == "ShExML"), ]
datasetShExML$Size <- factor(datasetShExML$Size)
datasetShExML$ShapesLanguage <- factor(datasetShExML$ShapesLanguage)

model = art(ElapsedTime_ms ~ ShapesLanguage + (1 | Size), data=datasetShExML)
marginal = art.con(model, "ShapesLanguage", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `F` * `Df` / (`F` * `Df` + `Df.res`))
Result
marginal

lm = artlm(model, "ShapesLanguage")
contrasts = summary(pairs(emmeans(lm, ~ ShapesLanguage)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Running the ART tool including post hoc tests and effect size for ShExML
```{r}
model = art(ElapsedTime_ms ~ ShapesLanguage + Size + ShapesLanguage:Size, data=datasetShExML)
marginal = art.con(model, "ShapesLanguage:Size", adjust="tukey")

Result = anova(model)
Result$part.eta.sq = with(Result, `Sum Sq`/(`Sum Sq` + `Sum Sq.res`))
Result

lm = artlm(model, "ShapesLanguage:Size")
contrasts = summary(pairs(emmeans(lm, ~ ShapesLanguage:Size)))

contrasts$d = contrasts$estimate / sigmaHat(lm)
contrasts
```

# Comparing linear scale on ShExML
```{r}
ggline(datasetShExML, x = "Size",
      y = "ElapsedTime_ms",
      color = "ShapesLanguage",
      add = "mean_sd")
```

# Comparing logarithmic scale on ShExML
```{r}
ggline(datasetShExML, x = "Size",
      y = "ElapsedTime_ms",
      ylab = "Elapsed Time (ms)",
      yscale = "log10",
      color = "ShapesLanguage",
      add = "mean_sd")
```